---
title: "Time Series Data"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr) # remotes::install_github("rstudio/learnr", build_vignettes = TRUE)
#library(gradethis) # remotes::install_github("rstudio/gradethis", build_vignettes = TRUE)

knitr::opts_chunk$set(echo = TRUE, comment = "")
#gradethis_setup()
#tutorial_options(exercise.reveal_solution = TRUE)

# https://bookdown.org/yihui/rmarkdown-cookbook/reuse-chunks.html#embed-chunk
<<user-facing-setup>>
```


```{r user-facing-setup, include=FALSE}
library(tidyverse)
library(tsibble)
library(lubridate)
library(fpp3)
theme_set(theme_bw())

<<def-weather1>>
<<as_tsibble-solution>>
<<def-daily_weather>>
<<def-weather3>>
<<def-weather4>>
```


## Examples

Here are a few examples of time-series data we might work with:

```{r}
simple_example <- tsibble(
  Year = 2015:2020,
  Observation = c(59, 20, 91, 44, 56, 12),
  index = Year)
simple_example
```

```{r}
olympic_running
```

```{r}
ansett
```



```{r}
PBS
```



##  `tsibbles`: the Index

A time series is a collection of data indexed by time. For example, let's look at some weather data:

```{r def-weather1}
weather1 <- nycflights13::weather %>%
  filter(origin == "EWR") %>% 
  select(origin, time_hour, temp)
```


```{r}
weather1
```

This is a `tibble`, which is adequate for storing time series data. But `t`**`s`**`ibble` (*time series* tibble) gives us some useful extra features.

We convert a `tibble` into a `tsibble` by specifying which variable is the `index`

```{r as_tsibble, exercise=TRUE}
weather2 <- as_tsibble(weather1, index = ____)
weather2
```

```{r as_tsibble-solution, exercise.eval=TRUE}
weather2 <- as_tsibble(weather1, index = time_hour)
weather2
```

We can now easily plot that:

```{r}
autoplot(weather2)
```

Notice how the x axis label shows that the measurements are taken every *hour* (`[1h]`). *What if we want the daily highs and lows?*

## Aggregating: Daily Highs and Lows

Let's look at an example timestamp.

```{r}
example_time <- ymd_hms("2022-03-14 01:59:26")
example_time
```

If we want to look at everything that happens on that day, we need to identify which *date* that timestamp falls on:

```{r}
as_date(example_time)
```

So new we can re-*index* by that date. Note two *syntax quirks*: we need to use ` ~ as_date` not just `as_date`, and we use `.` to refer to the *existing* index:

```{r def-daily_weather}
daily_weather <- weather2 %>% 
  index_by(date = ~ as_date(.)) %>% 
  summarize(
    temp_high = max(temp, na.rm = TRUE) # ignore missing values
  )
daily_weather
```

Exercise: plot the highest temperatures every *month* (use `yearmonth` to extract months, then `autoplot`).

```{r plot-high-temp, exercise=TRUE}
weather2 %>% 
  index_by(date = ~ as_date(.)) %>% 
  summarize(
    temp_high = max(temp, na.rm = TRUE)
  )
```

```{r plot-high-temp-solution}
# Could do ...
weather2 %>% 
  index_by(month = ~ yearmonth(.)) %>% 
  summarize(
    temp_high = max(temp, na.rm = TRUE)
  ) %>% 
  autoplot(.vars = vars(temp_high))

# or:
daily_weather %>% 
  index_by(month = ~ yearmonth(.)) %>% 
  summarize(
    temp_high = max(temp_high, na.rm = TRUE)
  ) %>% 
  autoplot(.vars = vars(temp_high))
```


```{r include=FALSE}
weather2 %>% 
  index_by(date = ~ as_date(.)) %>% 
  summarize(
    temp_high = max(temp),
    temp_low = min(temp)
  ) %>% 
#  autoplot(.vars = vars(temp_high, temp_low))
  pivot_longer(cols = c(temp_high, temp_low)) %>% 
  autoplot()
```

<!-- FIXME: Maybe a different exercise to avoid the pivot_longer? -->

## Sampling intervals

<!-- FIXME: is that term correct? -->

Time series can be sampled at different intervals. The weather data is hourly:

```{r}
tsibble::interval(weather2)
```

The daily weather was every 1 day:

```{r}
tsibble::interval(daily_weather)
```

Monthly data was every month:

```{r}
weather2 %>% index_by(month = ~ yearmonth(.)) %>% tsibble::interval()
```

But Olympic results are every 4 years:

```{r}
olympic_running %>% tsibble::interval()
```


To represent intervals of time, `tsibble` provides additional types. It uses standard types for days and sub-day units, though; you can use `lubridate` methods to make those:

```{r}
tsibble::yearmonth("2019 Jan") %>% class()
lubridate::ymd("2019-01-01") %>% class()
```


## Multiple variables in a time series

Weather stations give us not just temperature but other important factors too. We can store these as a time series, but now with several variables.

```{r def-weather3}
weather3 <- nycflights13::weather %>% 
  filter(origin == "EWR") %>% 
  select(time_hour, temp, humid, precip) %>% 
  as_tsibble(index = time_hour)
weather3
```

```{r}
# TODO EXERCISE: total precipitation each month?
```

## A `tsibble` is *tidy*

What about if we want to work with the weather in several different places? The *time* index will hopefully be the same, but there will be multiple *observations* at each time.

The `nycflights13::weather` data actually has three different airports:

```{r}
nycflights13::weather %>% count(origin)
```

If we try to make a `tsibble` out of this, we get an error:

```{r error=TRUE}
weather4 <- nycflights13::weather %>% 
  select(origin, time_hour, temp, humid, precip) %>% 
  as_tsibble(index = time_hour)
weather4
```

The problem is that there are times with multiple observations. We need to use a *key* to distinguish them.

```{r def-weather4}
weather4 <- nycflights13::weather %>% 
  select(origin, time_hour, temp, humid, precip) %>% 
  as_tsibble(index = time_hour,
             key = origin) #<< "origin" distinguishes the multiple observations at the same time.
weather4
```

The rule is that *each observation should be uniquely identified by index and key* (see the tsibble introductory vignette).

Now we can plot:

```{r}
weather4 %>% autoplot()
```

Often, plotting can reveal data issues. For example, it looks like there's an erroneous value in the JFK data. Let's go find it. *You may find this to be a helpful syntax reference for working with times*.

```{r}
weather4 %>% filter(temp < 15, time_hour > ymd("2013-04-01"), time_hour < ymd("2013-07-01"))
# Note: the `between` function doesn't work here. I don't know why. 
```

We can confirm that the value is incorrect by looking at [historical weather data](https://www.wunderground.com/history/daily/us/ny/new-york-city/KLGA/date/2013-5-8), or by comparing it with the temperature at other nearby airports:

*Note*: If I don't specify `tz`, I get a different hour. Why do you think that is?

```{r}
weather4 %>% filter(time_hour == ymd_hms("2013-05-08 22:00:00", tz = "America/New_York"))
```

You can play with this data here:

```{r play-with-weather4, exercise=TRUE}
weather4 %>% filter(time_hour == ymd_hms("2013-05-08 22:00:00", tz = "UTC"))
```



## Implicit grouping

`tsibble`s are implicitly grouped by their time index.

```{r}
olympic_running
olympic_running %>% summarize(m = mean(Time))
olympic_running %>% group_by(Length) %>% summarize(m = mean(Time))
```


## Appendix

To run any code chunk from this tutorial in your own environment, use:

```{r ref.label="user-facing-setup", echo=TRUE, eval=FALSE}

```

